---
title: "Homework 6"
Author: "Adam Rosenfeld"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(modelr)
library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```


# Problem 1
```{r message=FALSE}
homicide_df = 
  read_csv("data/homicide-data.csv", na = c("", "NA", "Unknown")) %>% 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    victim_age = as.numeric(victim_age),
    resolution = case_when(
      disposition == "Closed without arrest" ~ 0,
      disposition == "Open/No arrest"        ~ 0,
      disposition == "Closed by arrest"      ~ 1)
  ) %>% 
  filter(
    victim_race %in% c("White", "Black"),
    city_state != "Tulsa, AL",
    city_state != "Dallas, TX",
    city_state != "Phoenix, AZ",
    city_state != "Kansas City, MO") %>% 
  select(city_state, resolution, victim_age, victim_race, victim_sex)
```


Starting with Baltimore

```{r message=FALSE}
baltimore_df =
  homicide_df %>% 
  filter(city_state == "Baltimore, MD")

baltimore_model = 
glm(resolution ~ victim_age + victim_race + victim_sex, 
    data = baltimore_df,
    family = binomial()) %>% 
  broom::tidy() %>% 
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error)
  ) %>% 
  select(term, OR, starts_with("CI")) %>% 
  knitr::kable(digits = 3)
```

GLM model across all cities

```{r message=FALSE}
models_results_df = 
  homicide_df %>% 
  nest(data = -city_state) %>% 
  mutate(
    models = 
      map(.x = data, ~glm(resolution ~ victim_age + victim_race + victim_sex, data = .x, family = binomial())),
    results = map(models, broom::tidy)
  ) %>% 
  select(city_state, results) %>% 
  unnest(results) %>% 
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error)
  ) %>% 
  select(city_state, term, OR, starts_with("CI"))
```

Putting the estimates into a plot
```{r message=FALSE}
estimate_plot = 
  models_results_df %>% 
    filter(term == "victim_raceWhite") %>% 
    mutate(city_state = fct_reorder(city_state, OR)) %>% 
    ggplot(aes(x = city_state, y = OR)) + 
    geom_point() + 
    geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper)) + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(
      x = "City, State")

estimate_plot
```

This plot shows cities in order of increasing odds that crimes will be resolved if the victim is white. In Boston, White victims have over 10 times the odds of having their cases resolved than Black victims- while there's a large confidence interval, the odds ratio is still significantly high. The city with the lowest odds ratio is Tampa. 

## Problem 2

Reading in birthweight data
```{r}
baby_df = 
  read_csv("./data/birthweight.csv") %>% 
  mutate(
    babysex = case_when(
      babysex == 1 ~ "male",
      babysex == 2 ~ "female"
    ),
    frace = case_when(
      frace == 1 ~ "White",
      frace == 2 ~ "Black",
      frace == 3 ~ "Asian",
      frace == 4 ~ "Puerto Rican",
      frace == 8 ~ "Other",
      frace == 9 ~ "Unknown"
    ),
     mrace = case_when(
      mrace == 1 ~ "White",
      mrace == 2 ~ "Black",
      mrace == 3 ~ "Asian",
      mrace == 4 ~ "Puerto Rican",
      mrace == 8 ~ "Other"
    ),
    malform = case_when(
      malform == 0 ~ "absent",
      malform == 1 ~ "present"
    ))
```

Creating an initial model 
```{r message=FALSE}
model_test = lm(bwt ~ gaweeks, data = baby_df)
```

Making a prediction vs residual plot
```{r message=FALSE}
pred_resid_plot = 
  baby_df %>% 
    add_predictions(model_test) %>%
    add_residuals(model_test) %>% 
    ggplot(aes(x = pred, y = resid)) + 
    geom_point() +
    labs(
      x = "Predicted Birthweights",
      y = "Residuals")
  
pred_resid_plot
```

Cross-validation
```{r}
cv_df =
  crossv_mc(baby_df, 100, test = 0.2, id = "model_id") %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))

cv_df = 
  cv_df %>% 
  mutate(
    model_test_1  = map(train, ~lm(bwt ~ gaweeks, data = .x)),
    model_test_2 = map(train, ~lm(bwt ~  blength + gaweeks, data = .x)),
    model_test_3  = map(train, ~lm(bwt ~ bhead * blength * babysex, data = .x))) %>% 
  mutate(
    rmse_test_1 = map2_dbl(model_test_1, test, ~rmse(model = .x, data = .y)),
    rmse_test_2 = map2_dbl(model_test_2, test, ~rmse(model = .x, data = .y)),
    rmse_test_3 = map2_dbl(model_test_3, test, ~rmse(model = .x, data = .y)))

rmse_plot = 
  cv_df %>% 
    select(starts_with("rmse")) %>% 
    pivot_longer(
      everything(),
      names_to = "model", 
      values_to = "rmse",
      names_prefix = "rmse_") %>% 
    mutate(
      model = fct_inorder(model)) %>% 
    ggplot(aes(x = model, y = rmse)) + 
    geom_violin()

rmse_plot
```

# Problem 3

Loading in the data
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

